Cognitive Context: Impact on the Perception of a Robot

### 2015-07-03 main.tex
Proof-reading pass

\usepackage{tikz}
\usetikzlibrary{calc,decorations.pathreplacing}
\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}

@@

Robotics researchers often tend to consider that anthropomorphism only describes
a static set of human-like features of a robot (like its shape, its speech
capabilities, facial expressions, etc.).
Following~\citet{fink_anthropomorphism_2012}, we refer to these characteristics
as the \emph{anthropomorphic design} of the robot, and we call
\emph{anthropomorphism} the general \emph{social phenomenon} that emerges from
the (real or imagined) interaction between a non-human agent (a robot in our
case) and a human~\citep{persson_anthropomorphism_2000}.  According
to~\citet{epley_when_2008}, anthropomorphism includes the human perception of
emotional states, motivations, intentions in the non-human agent, and tends to
ascribe those qualities to it. As such, the dynamic and socio-cognitive
dimensions of anthropomorphism are essential to understand the complex bonds
that human users build with artificial agents such as
robots~\citep{lemaignan2014dynamics}. Therefore we propose therefore that
anthropomorphism can be a useful \emph{proxy} to these intricate psychological
bonds. This article aims at improving our understanding of this phenomenon
\textbf{by evidencing the impact of the cognitive presuppositions implied by the
context of the interaction}.

The study of context in HRI is not new. \textit{Context of use} is defined by
the ISO as \textit{``users, tasks, equipment (hardware, software and materials),
and the physical and social environments in which a product is used''}.
Accordingly, there are various aspects that can impact how a robot is
experienced. The real or imagined \textit{purpose} of a robot, including the
application context in which it is typically used (\eg office environment, home,
rehabilitation center, public space, \etc) as well as the task context (\eg
serious or playful task; composition of human-robot team; number of users) and
the (social) role in which the robot is used / experienced are taken into
account. Generally, when the context of use of a robot is social, entertaining
or playful, it will enhance anthropomorphism compared to when the context is a
routine or focused serious task (security, rescue, \etc).
\citet{joosse_what_2013} showed, for instance, that when the same robot ({\sc
nao}) was used in a different task context (cleaning task \vs tour guide), users
ascribed different personalities to the robot. \citet{goetz_cooperation_2002}
revealed a link between a robot's context of use and people's perception of the
robot: the authors found that people prefer a serious robot for serious tasks
and a less serious robot for more playful tasks. \citet{kaplan_free_2000}
discussed the role of uselessness in the design of robots. The author argues
that artificial pets like AIBO have no real purpose, in a sense that they do not
provide any kind of service or, in other words, \textit{``they are not doing
what you tell them to do''} and \textit{``they might refuse the order of its
owner''}. It may be this very aspect that increases people's tendency to
anthropomorphize the robot. According to Kaplan, it is in our daily use language
that we tend to attribute intentions to devices that are not doing their job
well.

@@

Measuring anthropomorphism raises
interesting methodological challenges, if only because people's perception of
robots change over the duration of their interaction with the
robots~\citep{lemaignan2014dynamics}. \citet{takayama_perspectives_2012}
discusses the effects of these changes and differentiates between what she calls
an \emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation whereas a \emph{reflective} perspective would
indicate one's reaction/sense of a situation based on a more in-depth
consideration and contemplation.  In the first phase of interaction with a
robot, people might respond ``mindlessly'' instead of responding
consciously~\citep{nass_machines_2000}, and only after what is called the
\emph{familiarization} period, one might respond in a more reflective manner
instead of an in-the-moment reaction: \citet{reeves_media_1996} showed for
instance that subjects interacting with technologies in ways close to how they
would behave towards people, would deny any human-like interaction in a post-hoc
questionnaire, illustrating this reflective perspective.

@@

\subsection*{Hypotheses}

The study has been designed out of the following intuition: \emph{the more we
ascribe cognitive capabilities to an agent, the more we focus on its head when
interacting with it}.  Eye-tracking was a natural methodology to investigate
this intuition, and we crafted audio-visual stimuli that would elicit different
levels of cognitive ascription onto the robot while being identical in every
other way.

We actually build several hypotheses: our baseline hypothesis \h{1} being that \emph{shallow
cognitive contexts} and \emph{deep cognitive contexts} are recognized
(manipulation check) and interpreted by the participants as being respectively
\emph{machine-like} (low level of anthropomorphic attributions) or
\emph{human-like} (higher level of anthropomorphic attributions).

Secondly, we hypothesize (\h{2}) that the \emph{difference} of gaze
behaviours between the human and robot conditions correlates with the
participants' initial tendency to anthropomorphize, measured from a standard
pre-questionnaire: we expect that people with high tendency to anthropomorphize
will tend to have smaller differences in their gaze behaviours while watching
either a robot or a human performing the same task.

Then, we hypothesize (\h{3}) that the gaze fixations on the head are significantly
different depending on the cognitive context, and that a deeper cognitive
context corresponds to longer gaze fixations on the head. This would in turn
interpreted at \emph{gaze fixations on the head is a proxy for an higher
level of human-likeness attribution}, the core hypothesis of this article.

Finally, the hypothesis \h{4} examines whether the cognitive context affects the
general tendency to anthropomorphize, \ie whether observing a
robot acting in an \emph{human-like} context leads to an increased level of
anthropomorphic attributions to robots (and vice versa). This also provides us
with a way to triangulate the behavioural measures with the subjective
attribution of anthropomorphism.

@@

As hinted, this study is build as an eye-tracking experiment (using a stationary
eye-tracker SMI RED 250) where participants watch short videos of an agent --
either a robot (Aldebaran Robotics' Nao) or a human -- performing very simple
task (picking an object or pointing toward a sound source), in two different
socio-cognitive contexts.  These contexts, created by priming and/or audio cues,
are designed to elicit either a shallow cognitive context (that participant
would refer as a \emph{machine-like situation}) or a deeper cognitive context
(refered as a \emph{human-like situation}).

Besides eye-tracking, the participants' perception of robots is assessed through a
pre- and a post-questionnaire interaction, described below.

@@

The study follows a 2$\times$2 design, summarized in table~\ref{table:design}. Our two
independent variables are the level of socio-cognitive context (\emph{shallow
cognitive context} \vs \emph{deep cognitive context}) and the nature of the
agent appearing in the stimulus (a \emph{robot}, eliciting a \emph{human-robot} interaction
situation, or a \emph{human}, eliciting a \emph{human-human} interaction
situation).

\begin{table}
    \centering
    \small

@@

The human actor was instructed to follow as closely as possible the actions and
attitudes of the robot (left/right glances, hesitations\ldots), while keeping a
natural, \emph{human-like} general behaviour.  Hence, the length of human videos
(57 seconds in total for the two tasks) was shorter than robot videos (110
seconds) due to the robot being usually slower in performing actions (in
particular walks) compared to human.

@@

Before starting the experiment, the participants were asked to fill a
pre-questionnaire which had 49 questions (5-point Likert scale). The
questionnaire comprised of questions assessing their familiarity with robots, 10
items from \emph{Big-Five} personality questionnaire, 15 questions taken from
Godspeed questionnaire (questions related to \emph{anthropomorphism},
\emph{likeability} and \emph{perceived intelligence}) and 20 questions rating
human-likeness (from~\cite{ruijten_introducing_2014})\footnote{The
the pre- and post-questionnaires are available online:
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking}.}. The pre-questionnaire
typically took 5 minutes for participants to fill. The results were turned into
a score noted \anti (\emph{initial tendency to anthropomorphize}) following a
method presented below (section~\ref{questionnaires_processing}).

The post-questionnaire was administered at the end of the experiment, and was
close to the pre-questionnaire: the personality questions were omitted, and a
manipulation check question was added (\emph{``In your opinion, the tasks that
you watched were more: A robot kind of task ... A human kind of task''}).
This questionnaire typically took 3 to 4 minutes to fill. The resulting score is
noted \antf (\emph{final tendency to anthropomorphism}).

The difference \deltaant measured for each participant between the pre-
and post-questionnaires forms the dependent variable of our experiment. It
reflects the impact of the video stimuli on the anthropomorphic perception that
they have of robots \emph{in general}: the Godspeed questions used in the
questionnaires are not specific to the Nao robot used in this particular
experiment, and indeed assess how humans perceive robots in a broad sense.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.3\linewidth]{experimentFlow}
    \caption{Overview of the course of the study}
    \label{course_of_study}
\end{figure}
\paragraph{Participants}

@@

\paragraph{Pre-questionnaire and familiarization}

Before the experiment, participants were made to read, understand and sign a
consent form. Participants were then asked to fill the 49 questions of
the pre-questionnaire, with a 10 minutes time limit that none of the participants
reached.

After the pre-questionnaire, we invited the participants to an initial
free interaction with a \textbf{powered-off} Nao robot (we did not constraint
the time, and participants played between 2 and 3 minutes with the robot).

The purpose of this pre-interaction was to get the participants acquainted with the
robot shape and kind of motions it could possibly perform, and therefore
mitigate the (visual) novelty effect (so that they do not get surprised seeing
a robot for the first time in the videos).

@@

\paragraph{Areas of interest}

@@

We recorded the fixations on one of these 10 zones and normalized by different total video duration and surface of AOIs between the robot and human conditions. We finally analyzed their distribution, producing what we
call \emph{gaze patterns}. The difference in the gaze patterns over ``N'' AOis can be calculated as follows:\\

\paragraph{Differences between gaze behaviours}

To compare gazing behaviours across conditions, we compute a \emph{difference in
gaze patterns} over the $N(=10)$ AOIs:

@@

Based on the experiments conducted on 55 participants in the human
($\mathcal{H}$) and robot ($\mathcal{R}$) interactions across \emph{shallow} and
\emph{deep cognitive context} conditions, we have the following
results.\footnote{The
detailed statistics and code used for obtaining the results are available here :
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking}.}

@@

While testing for all the 4 hypotheses, we observed no significant bias with
respect to age, gender, EPFL status, familiarity with robot and robot ownership
of the participants, \ie there was no correlation found between these factors
and the gaze patterns in the videos.

@@

We got a significant correlation between $\delta_{\mathcal{H},\mathcal{R}}$ and
the \anti. As can be seen in figure~\ref{h2}, there is a significant negative
correlation between the two quantities (Pearson Correlation Coefficient = -0.43,
$p < .001$).

\begin{figure}
    \centering
    \includegraphics[width=3.3in]{H2}\label{GazeDifference-vs-ICA}
    \caption{Difference in gaze patterns $\delta_{\mathcal{H},
    \mathcal{R}}^{\text{gaze}}$ \vs \anti}
    \label{h2}
\end{figure}

@@

We analysed these possible correlations for robot condition only: the human case is
inherently a source of deep cognitive context, hiding the effects of our
manipulation on a shallow \vs deep cognitive context.

Based on ANOVA results, taking the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf
hands}, {\sf torso}, {\sf legs} for the robot in both the shallow and deep
cognitive context conditions, the fixations on the head in the deep condition
were found to be significantly higher than that in the shallow condition
($F[1,54] = 4.2, p < .05$). For all other AOIs groups, the fixations across
shallow and deep conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOIs for the two
conditions.

\begin{figure}
    \centering
    \includegraphics[width=3.3in]{GazeHighLow}\label{GazeHighLow}
    \caption{Gaze fixations on the different areas on interest, for shallow \vs
    deep cognitive context conditions.}
    \label{h3}
\end{figure}

This constitutes the main result of this study: inducing a deeper cognitive
context for the same interaction suffices to lead to significantly longer
fixations on the agent's head.

@@

Running an ANOVA against the evolution of anthropomorphic attributions
(\deltaant) over the duration of the experiment, in the two cognitive
conditions, shows that \deltaant is greater for tasks with a deeper cognitive
context than that for the shallow cognitive tasks ($F[1,54] = 6.54, p < .05$,
figure~\ref{h4}). In other words, after being exposed to short videos of a robot
performing simple tasks, subjects primed with a deeper cognitive context report
a significantly higher tendency to anthropomorphize robots \emph{in general} (as
the questionnaires on anthropomorphic attributions explicitely refered to
\emph{robots in general}).

@@

The clearly significant correlation between deep \vs shallow conditions and the
answers to the question \emph{Were the task more robot-like or human-like?}
proves that our manipulation was successful, and that a shallow cognitive
context was indeed associated to a robot-like situation, while a deeper
cognitive context was associated to a human-like situation (\h{1}).

Hypothesis \h{2} looked at a possible interaction between our potential to
anthropomorphize (as measured in the pre-questionnaire) and our actual gaze
behaviour, the hypothesis being that high anthropomorphizers would show smaller
differences between the way they look at robots and humans (since they would
consider robots closer to humans than low anthropomorphizers). Our data supports
this hypothesis: participants reporting an higher tendency to anthropomorphize
tend to have similar gazing behaviours whether they look at humans or at robots.

To rephrase it: statistically speaking, your tendency to anthropomorphize things
is reflected in the way you look at these things: you will also look at them
\emph{as if they were humans}.

The third hypothesis \h{3} proposes that head fixations are related to the
cognitive context, which is clearly supported by our data: in deeper cognitive
context, one tends to spend more time looking at the head of the interacting
agent than in a shallow cognitive context. We also hypothesized that, in turn,
head fixations could act as a \emph{proxy to anthropomorphic projections}: as
underlined in the previous section, our data also support this second
hypothesis.

@@

The study and results presented in this article lead to two main contributions:
new insights on the impact of the cognitive context of the interaction on the
perception of a robot by humans; a novel unbiased methodology based on
eye-tracking to measure anthropomorphic attributions during a running
human-robot interaction.

Our first contribution extends our understanding of the attribution of
human-like characteristics to robot by exploring the role of the interaction
context. As we show in the article, it appears that \textbf{even relatively
subtle context priming may lead to significantly different anthropomorphic
perceptions of the exact same robot, performing the exact same task}, as
confirmed both by distinct gaze patterns, and different reported perceptions in
questionnaires. To the best of our knowledge, this effect had not been previously
evidenced.

This leads to a second-order effect that we also evidence here: depending on the
cognitive context, the tendency to attribute human-like characteristics to robot
\textbf{evolves} in significantly different ways: after observing a robot in a
context that pre-suppose deeper cognitive capabilities, \textbf{people increase
their general tendency to anthropomorphize robots} compared to a shallow
cognitive context, even though the robot appearance and visual behaviour is the
exact same.

@@

We initially prepared three different stimuli: the \emph{Picking} task, the
\emph{Sound} task and a \emph{Dance} task.  During the \emph{Dance} task, the
agent was either asked to \emph{``Show some movements''} (\emph{shallow
cognitive context} condition) or to \emph{``Dance with the music''}
(\emph{deeper cognitive context} condition). While all the participants were
shown this stimuli as well, we decided to exclude it from the results as the
nature of the task (participants had to watch body movements) defeated our
gaze-based investigation method: the subjects expectedly paid attention to the
whole body of the robot or human (even distribution of the gazes on the body),
thus making specific gaze patterns on the head difficult to evidence.

@@

%\subsection{Context and Interaction Design}
%
%\fxwarning{discuss here how context can be used to shape the interaction}
%
%
%Anthropomorphism as a proxy for cognitive/affective bonds that are esthablishing
%between humans and robots.
%
%\fxnote{discuss extensions towards mobile eyetrackers}
%
\bibliographystyle{apacite}
\bibliography{biblio}

>>>>

\usepackage{tikz}
\usetikzlibrary{shapes,positioning,calc,decorations.pathreplacing}
\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}

@@

Robotics researchers often tend to consider that anthropomorphism only describes
a static set of human-like features of a robot (like its shape, its speech
capabilities, facial expressions, etc.).
Following~\citet{fink_anthropomorphism_2012}, we refer to these characteristics
as the \emph{anthropomorphic design} of the robot, and we call
\emph{anthropomorphism} the general \emph{social phenomenon} that emerges from
the (real or imagined) interaction between a non-human agent (a robot in our
case) and a human~\citep{persson_anthropomorphism_2000}.  According
to~\citet{epley_when_2008}, anthropomorphism includes the human perception of
emotional states, motivations, intentions in the non-human agent, and tends to
ascribe those qualities to it. As such, the dynamic and socio-cognitive
dimensions of anthropomorphism are essential to understand the complex bonds
that human users build with artificial agents such as
robots~\citep{lemaignan2014dynamics}. Therefore anthropomorphism appears to be a
useful \emph{proxy} to study these intricate psychological
bonds. Going in that direction, this article introduces \textbf{new ways to assess
anthropomorphic attributions} and provides in particular \textbf{new insights on
the impact of the cognitive context of the interaction on anthropomorphic
attributions}.

\subsection{Role of context in human-robot interaction}

The study of context in human-robot interaction (HRI) is not new.
\textit{Context of use} is defined by the ISO as \textit{``users, tasks,
equipment (hardware, software and materials), and the physical and social
environments in which a product is used''}.  Accordingly, there are various
aspects that can impact how a robot is experienced. The real or imagined
\textit{purpose} of a robot, including the application context in which it is
typically used (\eg office environment, home, rehabilitation center, public
space, \etc) as well as the task context (\eg serious or playful task;
composition of human-robot team; number of users) and the (social) role in which
the robot is used / experienced are to be taken into account.

Notably, when the context of use of a robot is social, entertaining or playful,
it will enhance anthropomorphism compared to when the context is a routine or
focused serious task (security, rescue, \etc).  \citet{joosse_what_2013} showed,
for instance, that when the same robot ({\sc nao}) was used in a different task
context (cleaning task \vs tour guide), users ascribed different personalities
to the robot. \citet{goetz_cooperation_2002} revealed a link between a robot's
context of use and people's perception of the robot: the authors found that
people prefer a serious robot for serious tasks and a less serious robot for
more playful tasks. \citet{kaplan_free_2000} discussed the role of uselessness
in the design of robots. The author argues that artificial pets like AIBO have
no real purpose, in a sense that they do not provide any kind of service or, in
other words, \textit{``they are not doing what you tell them to do''} and
\textit{``they might refuse the order of its owner''}. It may be this very
aspect that increases people's tendency to anthropomorphize the robot. According
to Kaplan, it is in our daily use language that we tend to attribute intentions
to devices that are not doing their job well.

@@

Measuring anthropomorphism raises
interesting methodological challenges, if only because people's perception of
robots change over the duration of their interaction with the
robots~\citep{lemaignan2014dynamics}. \citet{takayama_perspectives_2012}
discusses the effects of these changes and differentiates between what she calls
an \emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation whereas a \emph{reflective} perspective would
indicate one's reaction/sense of a situation based on a more in-depth
consideration and contemplation.  In the first phase of interaction with a
robot, people might respond ``mindlessly'' instead of responding
consciously~\citep{nass_machines_2000}, and only after what is called the
\emph{familiarization} period, one might respond in a more reflective manner
instead of an in-the-moment reaction: \citet{reeves_media_1996} showed for
instance that subjects interacting with technologies in ways close to how they
would behave towards people, would still deny any human-like interaction in a post-hoc
questionnaire, illustrating this reflective perspective.

@@

\subsection{Hypotheses}

The study has been designed out of the following intuition: \emph{the more we
ascribe cognitive capabilities to an agent, the more we focus on its head when
interacting with it}. Eye-tracking would be a natural methodology to investigate
this intuition, and we would then design audio-visual stimuli that would elicit different
levels of cognitive ascription onto the robot while being identical in every
other way.

We actually build several hypotheses: our baseline hypothesis \h{1} is that we
can actually create such stimuli that induce recognisable \emph{shallow
cognitive contexts} and \emph{deep cognitive contexts} (manipulation check); and
further that these contexts are interpreted by the participants as being
respectively \emph{machine-like} (low level of anthropomorphic attributions) or
\emph{human-like} (higher level of anthropomorphic attributions).

Secondly, we hypothesize (\h{2}) that the \emph{difference} of gaze
behaviours between the human and robot conditions correlates with the
participants' initial tendency to anthropomorphize, measured from a standard
pre-questionnaire: we expect that people with high tendency to anthropomorphize
will tend to have smaller differences in their gaze behaviours while watching
either a robot or a human performing the same task: if this can be observed, it
would also show that anthropomorphic attributions are related to gaze patterns.

Then, we hypothesize (\h{3}) that the gaze fixations on the head are
significantly different depending on the cognitive context, and that a deeper
cognitive context corresponds to longer gaze fixations on the head. This would
in turn interpreted at \emph{gaze fixations on the head is a proxy for the level
of human-likeness attribution}, and constitutes the core hypothesis of this
article.

Finally, the hypothesis \h{4} examines whether the cognitive context affects the
general tendency to anthropomorphize, \ie whether observing \emph{one}
robot acting in an \emph{human-like} context leads to an increased level of
anthropomorphic attributions to robots \emph{in general}. This also provides us
with a way to triangulate the behavioural measures with the subjective
attribution of anthropomorphism: if such an effect is evidences, we can also
conclude that the cognitive context impacts the attribution of human-likeness,
and this, in turns, supports the hypothesis \h{3} that head fixations would
reflect not only the cognitive context but also the level of anthropomorphic
attributions.

Figure~\ref{hyps} illustrate how these four hypotheses relate to \emph{a)}
cognitive context, \emph{b)} anthropomorphic attributions, and \emph{c)} gaze
patterns.

\begin{figure}[ht]
\centering

\resizebox{0.5\linewidth}{!}{%

\begin{tikzpicture}[
    >=latex,
    every edge/.style={->,draw, very thick},
    every node/.style={draw, circle, text centered, text width=2.7cm}]

    \node at (0,0) (ctxt) {Cognitive context};
    \node at (-3,-3) (anthro) {Anthropomorphic attributions};
    \node at (3,-3)[circle split, text width=2.5cm] (gaze) {\textit{\small Fixations \\on head} \nodepart{lower} \textit{\small Gaze
distribution \\differences}};
    \node [draw=none,anchor=east,left=-0.5cm of gaze.west] (gazecap) {\large Gaze:};

    \path[every node/.style={draw=none}]
            (ctxt) edge [bend right] node[left] {\h{1}, \h{4}} (anthro)
            (ctxt) edge [bend left] node[right] {\h{3}} (gaze.north)
            (anthro) edge [bend right] node[above] {\h{2}} (gaze.south);
\end{tikzpicture}
}

\caption{Hypotheses of interactions between the three main concepts studied in
the article. Arrows mean ``A impacts B''.}

    \label{hyps}
\end{figure}

@@

As hinted, this study is build as an eye-tracking experiment (using a stationary
eye-tracker SMI RED 250) where participants watch short videos of an agent --
either a robot (Aldebaran Robotics' Nao) or a human -- performing very simple
task (picking an object or pointing toward a sound source), in two different
socio-cognitive contexts.  These contexts, created by priming and/or audio cues,
are designed to elicit either a shallow cognitive context (that participants
would \textit{a posteriori} rate as a \emph{machine-like situation}) or a deeper
cognitive context (rated as a \emph{human-like situation}).

Besides eye-tracking, the participants' perception of robots is assessed through a
pre- and a post-questionnaire, described below.

@@

The study follows a 2$\times$2 design, summarized in table~\ref{table:design}. Our two
independent variables are the level of socio-cognitive context (\emph{shallow
cognitive context} \vs \emph{deep cognitive context}) and the nature of the
agent appearing in the stimulus (a robot, eliciting a \emph{human-robot} interaction
situation, or a human, eliciting a \emph{human-human} interaction
situation).

\begin{table}[ht!]
    \vspace{2em}
    \centering
    \small

@@

The human actor was instructed to follow as closely as possible the actions and
attitudes of the robot (left/right glances, hesitations, gestures), while keeping a
natural, \emph{human-like} general behaviour.  Hence, the length of human videos
(57 seconds in total for the two tasks) was shorter than robot videos (110
seconds) due to the robot being usually slower in performing actions (in
particular walks) compared to human.

@@

Before starting the experiment, the participants were asked to fill a
pre-questionnaire which had 49 questions (5-point Likert scale). The
questionnaire comprised of questions assessing their familiarity with robots, 10
items from the \emph{Big-Five} personality questionnaire, 15 questions taken from
Godspeed questionnaire (questions related to \emph{anthropomorphism},
\emph{likeability} and \emph{perceived intelligence}) and 20 questions rating
human-likeness (from~\cite{ruijten_introducing_2014})\footnote{The
pre- and post-questionnaires are available online:
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking}.}. The pre-questionnaire
typically took 5 minutes for participants to fill. The results were turned into
a score noted \anti (\emph{initial tendency to anthropomorphize}) following a
method presented below (section~\ref{questionnaires_processing}).

The post-questionnaire was administered at the end of the experiment, and was
identical to the pre-questionnaire, except that the personality questions were omitted and a
manipulation check question was added (\emph{``In your opinion, the tasks that
you watched were more: A robot kind of task ... A human kind of task''}, on a 5
points scale).
This questionnaire typically took 3 to 4 minutes to fill. The resulting score is
noted \antf (\emph{final tendency to anthropomorphism}).

\anti, \antf and the difference \deltaant measured for each participant between
the pre- and post-questionnaires form the dependent variables of our experiment.
\deltaant, in particular, reflects the impact of the video stimuli on the
anthropomorphic perception that they have of robots \emph{in general}: the
Godspeed questions used in the questionnaires are not specific to the Nao robot
used in this particular experiment, and indeed assess how humans perceive robots
in a broad sense.

@@

\begin{figure}[ht]
\centering

\begin{tikzpicture}[
    >=latex,
    every edge/.style={<-,draw, very thick},
    every node/.style={draw, text centered, text width=5cm}]

    \node (n1) at (0,0) {Consent form};
    \node (n2) [below=of n1] {Familiarization}
        edge (n1);
    \node (n3) [below=of n2] {Pre-questionnaire}
        edge (n2);
    \node (n4) [below=of n3] {Videos}
        edge (n3);
    \node (n5) [below=of n4] {Post-questionnaire}
        edge (n4);

\end{tikzpicture}
    \caption{Overview of the course of the study}
    \label{course_of_study}
\end{figure}
\paragraph{Participants}

@@

\paragraph{Familiarization and pre-questionnaire}

Before the experiment, participants were made to read, understand and sign a
consent form. Participants were then asked to fill the 49 questions of
the pre-questionnaire, with a 10 minutes time limit (that none of the participants
reached).

After the pre-questionnaire, we invited the participants to an initial
free interaction with a \textbf{powered-off} Nao robot (we did not constraint
the time, and participants played between 2 and 3 minutes with the robot).
The purpose of this pre-interaction was to get the participants acquainted with the
robot shape and kind of motions it could possibly perform, and therefore
mitigate the (visual) novelty effect (so that they do not get surprised seeing
a robot for the first time in the videos).

@@

@@

We recorded the fixations on each of these 10 zones and normalized by different
total video duration and surface of AOIs between the robot and human conditions.
We finally analyzed their distribution, producing what we call \emph{gaze
patterns}.

To compare gazing behaviours across the robot and human conditions, we then
computed a \emph{difference in gaze patterns} over the $N(=10)$ AOIs:

@@

Based on the experiments conducted on 56 participants in the human and robot
interactions across \emph{shallow} and \emph{deep cognitive context} conditions,
we have the following results.\footnote{The detailed statistics and code used
    for obtaining the results are available here :
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking}.}

@@

While testing for all the 4 hypotheses, we observed no significant bias with
respect to age, gender, familiarity with robot and robot ownership
of the participants, \ie there was no effects to be found between these factors
and the other variables of the experiment.

@@

We found a significant correlation between
$\delta_{\mathcal{H},\mathcal{R}}^{\text{gaze}}$ and
the \anti. As seen in figure~\ref{h2}, there is a significant negative
correlation between the two quantities (Pearson Correlation Coefficient = -0.43,
$p < .001$). The higher the (initial) tendency to anthropomorphize, the smaller
the difference in gazing behaviours between the robot and human conditions. This
supports hypothesis \h{2}.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{H2}\label{GazeDifference-vs-ICA}
    \caption{Difference in gaze patterns $\delta_{\mathcal{H},
    \mathcal{R}}^{\text{gaze}}$ \vs \anti}
    \label{h2}
\end{figure}

@@

Based on ANOVA results, taking the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf
hands}, {\sf torso} and {\sf legs} for the robot in both the shallow and deep
cognitive context conditions, the fixations on the head in the deep condition
were found to be significantly higher than that in the shallow condition
($F[1,54] = 4.2, p < .05$). For all other AOIs groups, the fixations across
shallow and deep conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.6\columnwidth]{GazeHighLow}\label{GazeHighLow}
    \caption{Gaze fixations on the different areas on interest, for shallow \vs
    deep cognitive context conditions.}
    \label{h3}
\end{figure}

This constitutes the main result of this study: inducing a deeper cognitive
context for the same interaction suffices to lead to significantly longer
fixations on the agent's head (hypothesis \h{3}).

Note that we only analysed these possible correlations for robot condition: in
the human condition, the presence of the human is intrinsically a source of deep
cognitive context, hiding the effects of our manipulation on a shallow \vs deep
cognitive context.

@@

Running an ANOVA against the change of anthropomorphic attributions
(\deltaant) over the duration of the experiment in the two cognitive
conditions, shows that \deltaant is greater for tasks with a deeper cognitive
context than that for the shallow cognitive tasks ($F[1,54] = 6.54, p < .05$,
figure~\ref{h4}). In other words, after being exposed to short videos of a robot
performing simple tasks, subjects primed with a deeper cognitive context report
a significantly higher tendency to anthropomorphize robots \emph{in general} (as
the questionnaire on anthropomorphic attributions explicitly referred to
\emph{robots in general}).

@@

The significant correlation between deep \vs shallow conditions and the
answers to the question \emph{Were the task more robot-like or human-like?}
shows that our manipulation was successful, and that a shallow cognitive
context was indeed associated to a robot-like situation, while a deeper
cognitive context was associated to a human-like situation (\h{1}). This is
already an interesting result: subtle changes in the cognitive contexts are
recognized and clearly interpreted in terms of human-likeness.

Hypothesis \h{2} looked at a possible interaction between our tendency to
anthropomorphize (as measured in the pre-questionnaire) and our actual gaze
behaviour, the hypothesis being that high anthropomorphizers would show smaller
differences between the way they look at robots and humans (since they would
consider robots closer to humans than low anthropomorphizers). Our data supports
this hypothesis: participants reporting an higher tendency to anthropomorphize
tend to have closer gazing behaviours when they look at humans or at robots:
statistically speaking, the tendency to anthropomorphize robots
is reflected in the way one looks at them: you look at them
like you would look at humans.

The third hypothesis \h{3} proposes that head fixations are related to the
cognitive context, which is clearly supported by our data: in deeper cognitive
context, one tends to spend more time looking at the head of the interacting
agent than in a shallow cognitive context. We also hypothesized that, in turn,
head fixations could act as a \emph{proxy to anthropomorphic projections}: as
underlined in the previous section, our data also support this.

@@

The study and results presented in this article lead to two main contributions:
new insights on the impact of the cognitive context on the perception of a robot
by humans; a novel unbiased methodology based on eye-tracking to assess
anthropomorphic attributions during a running human-robot interaction.

Our first contribution extends our understanding of the attribution of
human-like characteristics to robot by exploring the role of the interaction
context. As we show, it appears that \textbf{even relatively
subtle context priming may lead to significantly different anthropomorphic
perceptions of the exact same robot, performing the exact same task}, as
confirmed both by distinct gaze patterns, and different reported perceptions in
questionnaires.

This leads to a second-order effect that we also evidence here: depending on the
cognitive context, the tendency to attribute human-like characteristics to robot
\textbf{evolves} in significantly different ways. \textbf{After observing a robot in a
context that pre-suppose deeper cognitive capabilities, people increase
their general tendency to anthropomorphize robots} compared to a shallow
cognitive context, even though the robot appearance and visual behaviour is the
exact same.

@@

We initially prepared three different stimuli: the \emph{Picking} task, the
\emph{Sound} task and a \emph{Dance} task.  During the \emph{Dance} task, the
agent was either asked to \emph{``Show some movements''} (\emph{shallow
cognitive context} condition) or to \emph{``Dance with the music''}
(\emph{deeper cognitive context} condition). While all the participants were
shown this stimuli as well, we decided to exclude it from the results as the
nature of the task (participants had to watch body movements) defeated our
gaze-based investigation method: the subjects expectedly paid attention to each
and every part of the body of the robot or human (even distribution of the gazes
on the body), thus making specific gaze patterns on the head difficult to
evidence.

@@

\section*{Acknowledgments}

We would like here to acknowledge the contribution of Julia Fink to a better
understanding of anthropomorphism as a psychological phenomenon.

\bibliographystyle{apacite}
\bibliography{biblio}
